Miniproject - Combinatorics

### 2021-12-14 Combinatorics/submission.tex
added rust syntax highlighting to combinatorics
Fixed alg geo, hopefully finally

\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage[style=english]{csquotes}
\usepackage[language=english, backend=biber, style=alphabetic, sorting=nyt]{biblatex}

@@

\section{Part I}

\begin{proposition}
    \label{prop:definition_rank_function}
    Let $P$ be a graded poset in which every maximal chain has length $n + 1$. Then the function
    \begin{equation*}
        r: P \to \{ 0, ..., n \}, \quad x \mapsto \max\{ k \in \N \ | \ \exists a_1, ..., a_k \in P: \ a_1 < a_2 < ... < a_k < x \}
    \end{equation*}
    is well-defined and the unique function with $x < y$ implies $r(x) < r(y)$ for all $x, y \in P$.
    We call it the \emph{rank function} of $P$.
\end{proposition}
\begin{proof}
    Clearly $r$ is well-defined, as for every increasing sequence $a_1 < ... < a_k < x$, we have the chain $C := \{ a_1, ..., a_k, x \}$ of size $k + 1$.
    Hence by assumption, $k \leq n$ and so $r(x)$ is finite and in $\{ 0, ..., n \}$.

@@

    Hence we find $r(b_i) < r(b_{i + 1})$ and so inductively that $r(b_i) \geq i - 1$ and $r(b_i) \leq i - 1$ since $r(b_1) \geq 0$ and $r(b_{n + 1}) \leq n$.
    It follows that $r(b_i) = i - 1$ and so $x = b_{r(x) + 1}, y = {r(y) + 1}$.
    Therefore, we have a chain
    \begin{equation*}
        x < b_{r(x) + 2} < ... < b_{r(y)} < y
    \end{equation*}
    of length $r(y) - r(x) - 1$.
    Statement (ii) follows directly, as in this setting, have $r(y) \geq r(x) + 2$ and so $r(y) - r(x) - 1 \neq 0$.
\end{proof}

\section{Part II}
\begin{proposition}
    \label{prop:characterization_lym}
    For a graded poset $P$ with layers $L_0, ..., L_n$ the following statements are equivalent:
    \begin{itemize}
        \item For every antichain $A \subseteq P$ have
        \begin{equation*}
            \sum_{i = 0}^n \frac {|A \cap L_i|} {|L_i|} \leq 1
        \end{equation*}
        \item For each $1 < i \leq n$ and $F \subseteq L_i$ have
        \begin{equation*}
            \frac {|\partial F|} {|L_{i - 1}|} \geq \frac {|F|} {|L_i|}
        \end{equation*}
        where
        \begin{equation*}
            \partial F := \{ a \in L_{i - 1} \ | \ \exists b \in F: \ a \leq b \}
        \end{equation*}
        \item There exists a (nonempty) sequence of maximal chains $C_1, ..., C_t$ such that for all $1 \leq i \leq n$ we have:
        \begin{equation*}
            |\{ j \ | \ x \in C_j \}| = |\{ j \ | \ y \in C_j \}| \ \text{for all $x, y \in L_i$}
        \end{equation*}
    \end{itemize}
\end{proposition}
\begin{proof}
    Show the directions (ii) $\Rightarrow$ (iii) $\Rightarrow$ (i) $\Rightarrow$ (ii)
    \paragraph{(ii) $\Rightarrow$ (iii)} Use induction on $n$.
    Again, the base case is trivial, just take chains $\{x\}$ for each $x \in A = L_0$.
    So assume that $n > 0$. The set
    \begin{equation*}
        A' := \bigcup_{i = 0}^{n - 1} L_i
    \end{equation*}
    is a graded poset of maximal rank $n - 1$, and hence there are maximal chains $C_1, ..., C_t \subseteq A'$ such that for all $1 \leq i \leq n - 1$ have
    \begin{equation*}
        |\{j \ | \ x \in C_j\}| = |\{j \ | \ y \in C_j\}| \ \text{for all $x, y \in L_i$}
    \end{equation*}
    by induction hypothesis.
    Let now $V_1, V_2 := [t] \times L_n$ (treat them as disjoint) and consider the bipartite graph $G := (V_1 \sqcup V_2, E)$ where $E$ is defined as follows:
    \begin{equation*}
        \{ \underbrace{(i, a)}_{\in V_1}, \underbrace{(j, b)}_{\in V_2} \} \in E \ \Leftrightarrow \ \max C_j < a
    \end{equation*}
    We use Hall's theorem to show that $G$ has a perfect matching.

    Consider any $W = \{ (i_1, a_1), ..., (i_w, a_w) \} \subseteq V_1$ and let $F = \{ a_1, ..., a_w \}$.
    Then
    \begin{equation*}
        W \subseteq [t] \times F \ \Rightarrow \ |W| \leq t|F|
    \end{equation*}
    By choice of $C_1, ..., C_t$, we have that the number of $j$ with $x \in C_j$ is the same for all $x \in L_{n - 1}$, say $k$.
    Since the $C_i$ are maximal chains, each intersects $L_{n - 1}$ in exactly one element.
    So have bijection
    \begin{align*}
        \{ (x, i) \ | \ x \in L_{n - 1}, 1 \leq i \leq t, x \in C_i\} \to \{ i \ | \ 1 \leq i \leq t \}, \quad (x, i) \mapsto i
    \end{align*}
    where the set on the left-hand side has size $k |L_{n - 1}|$. It follows that $k|L_{n - 1}| = t$.

    Since $\max C_j \in L_{n - 1}$ for all $j$, we have
    \begin{align*}
        N(W) =& \{ (j, b) \ | \ b \in L_n, \exists a \in F: \max C_j < a \} \\
        =& L_n \times \{ j \ | \ \exists a \in F: \max C_j < a \} \\
        =& L_n \times \{ j \ | \ \max C_j \in \partial F \}
    \end{align*}
    and so by the above
    \begin{align*}
        |N(W)| = |L_n| \sum_{\max C_j \in \partial F} 1 = |L_n| \sum_{x \in \partial F} \sum_{x \in C_j} 1 = |L_n| \sum_{x \in \partial F} k = |L_n| |\partial F| k
    \end{align*}
    Together with the assumption that $|\partial F| |L_n| \geq |F| |L_{n - 1}|$ we see that
    \begin{equation*}
        |W| \leq t |F| = k |L_{n - 1}| |F| \leq k |\partial F| |L_n| = |N(W)|
    \end{equation*}
    So Hall's theorem yields a perfect matching $M \subseteq E$ from $V_1$ to $V_2$.
    As $|V_1| = |V_2|$, this is already a 1-to-one correspondence.

    Now consider the sets
    \begin{equation*}
        C'_m := C_j \cup \{ a \} \ \text{where} \ m = \{\underbrace{(i, a)}_{\in V_1}, \underbrace{(j, b)}_{\in V_2}\} \in M
    \end{equation*}
    These are chains, as $\max C_j < a$ for each $C'_m$.
    Further, for each $x, y \in L_i, i < n$ have that
    \begin{equation*}
        |\{ C'_m \ | \ x \in C'_m \}| = |L_n \times \{ j \ | \ x \in C_j \}| = |L_n \times \{ j \ | \ y \in C_j \}| = |\{ C'_m \ | \ y \in C'_m \}|
    \end{equation*}
    as $M$ is a matching from $V_2$ to $V_1$.
    Finally, for all $x \in L_n$ we have that
    \begin{align*}
        |\{ C'_m \ | \ x \in C'_m \}| =& |\{ m \in M \ | \ \exists i, v \in V_2: m = \{ (i, x), v \} \}| = |\{ (i, x) \ | \ (i, x) \in V_1 \}| = t
    \end{align*}
    as $M$ is a matching from $V_1$ to $V_2$.

@@

After all dependencies are present, the following code will yield the result.
\lstinputlisting{./src/main.rs}

>>>>

\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage{listings-rust}
\usepackage[style=english]{csquotes}
\usepackage[language=english, backend=biber, style=alphabetic, sorting=nyt]{biblatex}

@@

\section{Part I}
First of all, we introduce the notion of a rank function on a graded poset.
\begin{proposition}
    \label{prop:definition_rank_function}
    Let $P$ be a graded poset in which every maximal chain has length $n + 1$. Then the function
    \begin{equation*}
        r: P \to \{ 0, ..., n \}, \quad x \mapsto \max\{ k \in \N \ | \ \exists a_1, ..., a_k \in P: \ a_1 < a_2 < ... < a_k < x \}
    \end{equation*}
    is well-defined and the unique function such that $x < y$ implies $r(x) < r(y)$ for all $x, y \in P$.
    We call it the \emph{rank function} of $P$.
\end{proposition}
\begin{proof}
    Clearly $r$ is well-defined, as for every increasing sequence $a_1 < ... < a_k < x$, we have the chain $C := \{ a_1, ..., a_k, x \}$ of size $k + 1$.
    Hence by assumption, $k \leq n$ and so $r(x)$ is finite and in $\{ 0, ..., n \}$.

@@

    Hence we find $r(b_i) < r(b_{i + 1})$ and so inductively that $r(b_i) \geq i - 1$ and $r(b_i) \leq i - 1$ since $r(b_1) \geq 0$ and $r(b_{n + 1}) \leq n$.
    It follows that $r(b_i) = i - 1$ and so $x = b_{r(x) + 1}, y = b_{r(y) + 1}$.
    Therefore, we have a chain
    \begin{equation*}
        x < b_{r(x) + 2} < ... < b_{r(y)} < y
    \end{equation*}
    of length $r(y) - r(x) - 1$.
    Statement (ii) follows directly, as in this setting, have $r(y) \geq r(x) + 2$ and so $r(y) - r(x) - 1 \neq 0$.
\end{proof}

\section{Part II}
In this part, we want to study whether graded posets satisfy an analogue of the LYM inequality.
\begin{proposition}
    \label{prop:characterization_lym}
    For a graded poset $P$ with layers $L_0, ..., L_n$ the following statements are equivalent:
    \begin{itemize}
        \item For every antichain $A \subseteq P$ have
        \begin{equation*}
            \sum_{i = 0}^n \frac {|A \cap L_i|} {|L_i|} \leq 1
        \end{equation*}
        \item For each $1 < i \leq n$ and $F \subseteq L_i$ have
        \begin{equation*}
            \frac {|\partial F|} {|L_{i - 1}|} \geq \frac {|F|} {|L_i|}
        \end{equation*}
        where
        \begin{equation*}
            \partial F := \{ a \in L_{i - 1} \ | \ \exists b \in F: \ a \leq b \}
        \end{equation*}
        \item There exists a (nonempty) sequence of maximal chains $C_1, ..., C_t$ such that for all $1 \leq i \leq n$ we have:
        \begin{equation*}
            |\{ j \ | \ x \in C_j \}| = |\{ j \ | \ y \in C_j \}| \ \text{for all $x, y \in L_i$}
        \end{equation*}
    \end{itemize}
\end{proposition}
\begin{proof}
    Show the directions (ii) $\Rightarrow$ (iii) $\Rightarrow$ (i) $\Rightarrow$ (ii)
    \paragraph{(ii) $\Rightarrow$ (iii)} Use induction on $n$.
    The base case is trivial, just take chains $\{x\}$ for each $x \in A = L_0$.
    So assume that $n > 0$. The set
    \begin{equation*}
        A' := \bigcup_{i = 0}^{n - 1} L_i
    \end{equation*}
    is a graded poset of maximal rank $n - 1$, and hence there are maximal chains $C_1, ..., C_t \subseteq A'$ such that for all $1 \leq i \leq n - 1$ have
    \begin{equation*}
        |\{j \ | \ x \in C_j\}| = |\{j \ | \ y \in C_j\}| \ \text{for all $x, y \in L_i$}
    \end{equation*}
    by induction hypothesis.
    Let now $V_1 := \{1\} \times [t] \times L_n$ and $V_2 := \{2\} \times [t] \times L_n$ and consider the bipartite graph $G := (V_1 \sqcup V_2, E)$ where $E$ is defined as follows:
    \begin{equation*}
        \{ \underbrace{(1, i, a)}_{\in V_1}, \underbrace{(2, j, b)}_{\in V_2} \} \in E \ \Leftrightarrow \ \max C_j < a
    \end{equation*}
    We use Hall's theorem to show that $G$ has a matching.

    Consider any $W = \{ (1, i_1, a_1), ..., (1, i_w, a_w) \} \subseteq V_1$ and let $F = \{ a_1, ..., a_w \}$.
    Then
    \begin{equation*}
        W \subseteq [t] \times F \ \Rightarrow \ |W| \leq t|F|
    \end{equation*}
    By choice of $C_1, ..., C_t$, we have that the number of $j$ with $x \in C_j$ is the same for all $x \in L_{n - 1}$, say $k$.
    Since the $C_i$ are maximal chains, each intersects $L_{n - 1}$ in exactly one element.
    So have bijection
    \begin{align*}
        \{ (x, i) \ | \ x \in L_{n - 1}, 1 \leq i \leq t, x \in C_i\} \to \{ i \ | \ 1 \leq i \leq t \}, \quad (x, i) \mapsto i
    \end{align*}
    where the set on the left-hand side has size $k |L_{n - 1}|$. It follows that $k|L_{n - 1}| = t$.

    Since $\max C_j \in L_{n - 1}$ for all $j$, we have
    \begin{align*}
        N(W) =& \{ (2, j, b) \ | \ b \in L_n, \exists a \in F: \max C_j < a \} \\
        =& \{2\} \times \{ j \ | \ \exists a \in F: \max C_j < a \} \times L_n \\
        =& \{2\} \times \{ j \ | \ \max C_j \in \partial F \} \times L_n
    \end{align*}
    and so by the above
    \begin{align*}
        |N(W)| = |L_n| \sum_{\max C_j \in \partial F} 1 = |L_n| \sum_{x \in \partial F} \sum_{x \in C_j} 1 = |L_n| \sum_{x \in \partial F} k = |L_n| |\partial F| k
    \end{align*}
    Together with the assumption that $|\partial F| |L_n| \geq |F| |L_{n - 1}|$ we see that
    \begin{equation*}
        |W| \leq t |F| = k |L_{n - 1}| |F| \leq k |\partial F| |L_n| = |N(W)|
    \end{equation*}
    So Hall's theorem yields a matching $M \subseteq E$ from $V_1$ to $V_2$.
    As $|V_1| = |V_2|$, this is already a 1-to-one correspondence.

    Now consider the sets
    \begin{equation*}
        C'_m := C_j \cup \{ a \} \ \text{where} \ m = \{\underbrace{(1, i, a)}_{\in V_1}, \underbrace{(2, j, b)}_{\in V_2}\} \in E
    \end{equation*}
    These are chains, as $\max C_j < a$ for each $C'_m$.
    Further, for each $x, y \in L_i, i < n$ have that
    \begin{equation*}
        |\{ C'_m \ | \ x \in C'_m \}| = |L_n \times \{ j \ | \ x \in C_j \}| = |L_n \times \{ j \ | \ y \in C_j \}| = |\{ C'_m \ | \ y \in C'_m \}|
    \end{equation*}
    as $M$ is a matching from $V_2$ to $V_1$.
    Finally, for all $x \in L_n$ we have that
    \begin{align*}
        |\{ C'_m \ | \ x \in C'_m \}| =& |\{ m \in M \ | \ \exists i, v \in V_2: m = \{ (i, x), v \} \}| = |\{ (i, x) \ | \ (i, x) \in V_1 \}| = t
    \end{align*}
    as $M$ is a matching from $V_1$ to $V_2$.

@@

After all dependencies are present, the following code will yield the result.
\lstinputlisting[language = rust]{./src/main.rs}
